Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-03-08 cluster_paper/ms.tex
rotation references: use Curtis+20 table 5

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars not having yet reached the
    main-sequence.
    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
     The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $Bp-Rp\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  However below this mass, the data and models diverge
at colors redder than Bp-Rp$\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Explanations invoking both starspots and
incomplete molecular line lists have been proposed \citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the PARSEC isochrones since they diverge at slightly lower
mass than the MIST models, due to the temperature-opacity calibration
implemented by \citet{chen_improving_2014}.  For purposes of
age-dating the cluster, we prefer to focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal predictive
power.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
they are: stars past the turnoff in both the halo and core are on a
consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
the precision with which the claim can be made.  The MSTO stars are
consistent with the 178 and 316 Myr models, but are bluer than the 100
Myr model.  The RGB stars are consistent only with the 178 Myr
($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
the width of the MSTO can be attributed to binary stars, we are most
interested in the blue edge. This appears most compatible with the 178
Myr model.  These results are consistent with the MSTO age range of
165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
work for a more precise model-dependent comparison.
This implies an absolute age slightly older than the Pleiades
\citep[{\it cf.}][]{mermilliod_comparative_1981}.

@@

\subsubsection{Cluster Star Sample}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3a.pdf}
			\includegraphics[width=0.49\textwidth]{f3b.pdf}
		}
	   %
		%\subfloat{
		%	\includegraphics[width=0.6\textwidth]{f3b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    Subset $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom})
    undergo successive degrees of automated cleaning, described in the
    text.
    The Pleiades \citep[125\,Myr][]{rebull_rotation_2016a} and
    Praesepe \citep[650\,Myr][]{douglas_poking_2017} are shown for
    reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at (Bp-Rp)$_0$$<$1.25.
    There is a narrow range of color, from 1.25$<$(Bp-Rp)$_0$$<$1.75,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

@@

Figure~\ref{fig:rot} shows the resulting rotation periods, following
three different sets of cleaning criteria.  The first and second sets
of criteria were entirely automated, and yielded subsets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For both subsets, we considered
light curves for which the peak Lomb-Scargle periodogram period was
below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
for which no equal-brightness or greater companions were within the
aperture.  We also required that the target star be the brightest star
in the aperture, and that at most one companion with brightness
exceeding one tenth of the target star's brightness be present in the
aperture according to the Gaia DR2 source catalog.  For subset
$\mathcal{A}$, this yielded \nautorotnumerator light curves, out of
\nautorotdenominator stars for which rotation periods might plausibly
have been detected.  For subset $\mathcal{B}$, we additionally
required that the Lomb-Scargle and Stellingwerf phase-dispersion
methods yielded periods identical to within 10\%.  This yielded
\nautorotnumeratormatching light curves.
Subset $\mathcal{A}$ is more complete than 
subset $\mathcal{B}$, at the expense of greater contamination.

{\bf TODO CORRECTLY}
To put the clusters on the same color scale, we needed to adopt
reddening corrections.
For \cn, we used the same average reddening as described in
Section~\ref{subsec:hr}.
% FIXME: where did these reference tables come from???
For the Pleiades, and for Praesepe, we used {\bf XXX, YYY}.

{\bf TODO CORRECTLY}
The resulting gyrochronology age we find for the core is XXX.
For the halo, the claimed age from gyrochronology is YYY.
Applying the same procedure to the field star comparison sample (see
Appendix~\ref{app:compstar}),
we get an age of ZZZ+/-AAA.

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars not having yet reached the
    main-sequence.
    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
     The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $Bp-Rp\approx$2.2).
    \label{fig:hr}
  }
\end{figure*}

@@

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  However below this mass, the data and models diverge
at colors redder than $Bp-Rp\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Explanations invoking both starspots and
incomplete molecular line lists have been proposed \citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the PARSEC isochrones since they diverge at slightly lower
mass than the MIST models, due to the temperature-opacity calibration
implemented by \citet{chen_improving_2014}.  For purposes of
age-dating the cluster, we prefer to focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal predictive
power.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
they are: stars past the turnoff in both the halo and core are on a
consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
the precision with which the claim can be made.  The MSTO stars are
consistent with the 178 and 316 Myr models, but are bluer than the 100
Myr model.  The RGB stars are consistent only with the 178 Myr
($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
the width of the MSTO can be attributed to binary stars, we are most
interested in the blue edge. This appears most compatible with the 178
Myr model.  These results are consistent with the MSTO age range of
165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
work for a more precise model-dependent comparison.
This implies an absolute age slightly older than the Pleiades
\citep[cf.][]{mermilliod_comparative_1981}.

@@

\subsubsection{Cluster Star Sample}
\label{subsubsec:cluster}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3a.pdf}
			\includegraphics[width=0.49\textwidth]{f3b.pdf}
		}
	   %
		%\subfloat{
		%	\includegraphics[width=0.6\textwidth]{f3b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The core and halo of \cn\ in the space of rotation
    period and dereddened Gaia color.}
    Subset $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom})
    undergo successive degrees of automated cleaning, described in Section~\ref{subsubsec:cluster}.
    The Pleiades \citep[120\,Myr;][]{rebull_rotation_2016a} and
    Praesepe \citep[650\,Myr;][]{douglas_poking_2017} are shown for
    reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$,
    in which some NGC\,2516 members appear elevated with respect to
    the Pleiades, suggesting a slightly older age.
		\label{fig:rot}
	}
\end{figure*}

@@

Figure~\ref{fig:rot} shows the resulting rotation periods, following
three different sets of cleaning criteria.  The first and second sets
of criteria were entirely automated, and yielded subsets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For both subsets, we considered
light curves for which the peak Lomb-Scargle periodogram period was
below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
for which no equal-brightness or greater companions were within the
aperture.  Beyond requiring that the target star be the brightest star
in the aperture, we also required that at most one companion with
brightness exceeding one tenth of the target star's brightness be
present in the aperture according to the Gaia DR2 source catalog.  For
set $\mathcal{A}$, this yielded \nautorotnumerator light curves,
out of \nautorotdenominator stars for which rotation periods might
plausibly have been detected.  For set $\mathcal{B}$, we
then additionally required that the Lomb-Scargle and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%.
This yielded \nautorotnumeratormatching light curves.

The comparison to the Pleiades and Praesepe is made using the rotation
periods and dereddened colors from Table~5 of \citet{curtis_rup147_2020}, which
were based data from \citet{rebull_rotation_2016a} and
\citet{douglas_k2_2019} respectively.

A few conclusions can be drawn by comparing
the \cn\ rotation periods with those of
the Pleiades and Praesepe.
The first is that in both stages of cleaning,
the core and halo of \cn\ both show
gyrochronological ``slow sequences''.
These sequences appear to be
coeval with each other, and at least to first order
appear to overlap with the Pleiades sequence.

Second, subset
$\mathcal{A}$ is more complete than subset $\mathcal{B}$.
This completeness comes at the
expense of greater contamination.
The slow sequence
has a factor of $\sim$twice as many stars in subset
$\mathcal{A}$ than $\mathcal{B}$.
However stars above the slow sequence, which could
well be field contaminants, for the most part disappear in
subset $\mathcal{B}$.
The ``void'' underneath the slow sequence
is also more pronounced in the cleaner
subset,
though it seems to not be as empty as in
the Pleiades.

Finally, comparing the slow sequences of the Pleaides
and \cn, they appear to overlap from roughly
$0.5<(Bp-Rp)_0<1.2$.
For stars that are redder,
the dispersion in rotation periods increases.
From
$1.2<(Bp-Rp)_0<1.7$, subset
$\mathcal{B}$ of \cn\ has stars that overlap with 
the Pleiades.
However, it also has about ten stars at longer rotation periods---extending
up to $\approx$11 days, rather than the
$\approx$8.5day limit for the Pleiades.
This is consistent with \cn\ being slightly
older than the Pleiades.

{\bf We explore this quantitatively by fitting
interpolated models}[?].
The resulting gyrochronology age we find for the core is XXX.
For the halo, the claimed age from gyrochronology is YYY.
Applying the same procedure to the field star comparison sample (see
Appendix~\ref{app:compstar}),
we get an age of ZZZ+/-AAA.
